### 2013-03-01 9_backmatter/appendix.tex
201303_Mar01

\chapter{Consistent Measures of Distance for Multivariate POLSAR Data} %chapter 4
\label{chap:polsar}

In this chapter, some statiscally consistent measures of distance for multi-variate POLSAR data is presented.
The first section introduces the basics of POLSAR statiscal analysis and the log-transformation proposed in this thesis.
The second section presents conclusive evidence that POLSAR data is multiplicative and heteroskedastic in the original domain.
Then log-transformation is shown converting these into an additive and homoskedastic model.
Consequently a few consistent measures of distance, namely the dispersion and the contrast measurements, is presented in the third section.
Section (TODO:REF) illustrates how a few state-of-the-art results in POLSAR processing might be explained using the present approach.
From another perspective, section (TODO:REF) discusses how the time-tested results for univariate SAR can also be considered as a special case for the multi-variate POLSAR results.
Finally the last section strengthens our model showing the matching between simulation and real-life captured data
  not only for a few dimensionally-collapsed single-dimensional measures
  but also for all elements of the multi-dimensional POLSAR covariance matrix.

\section{POLSAR statistical analysis}

In this section, the POLSAR scattering vector is denoted as $s$.
In the case of partial polarimetric SAR (single polarization in transmit and dual polarization in receipt),
  the vector is two-dimensional ($d=2$) and is normally written as: 
\begin{equation}
s_{part}=\begin{bmatrix}
S_h\\ 
S_v
\end{bmatrix}
\end{equation}
In the case of full and monostatic POLSAR data,
  the vector is three-dimensional ($d=3$) and is presented as:
\begin{equation}
s_{full}=\begin{bmatrix}
S_{hh}\\
\sqrt{2}S_{hv}\\
S_{vv}
\end{bmatrix}
\end{equation}

Let $\Sigma=E [ss^{*T}]$ denotes the population expected value of the POLSAR covariance matrix,
  where $s^{*T}$ denotes the complex conjugate transpose of $s$. 
Assuming all the elements in $s$ are independent
  and $s$ is jointly circular complex Gaussian with the given covariance matrix $\Sigma$,
  then the probably density function (PDF) of $s$ can be written as:
\begin{equation}
  pdf(s;\Sigma)=\frac{1}{\pi^d|\Sigma|} e^{-s^{*T}\Sigma^{-1}s}
\end{equation}
where $|M|$ denotes the determinant of the matrix $M$.

As covariance matrix is only defined on multiple data points,
  sample covariance matrix of POLSAR data is commonly presented in ``ensemble'' format.
They are formed as the mean of Hermitian outer product of single-look scattering vectors
\begin{equation}
  C_v = \langle ss^{*T} \rangle = \frac{1}{L} \sum^L_{i=1}s_is_i^{*T}
\end{equation}
with $s_i$ denotes the single-look scattering vector,
  which equals $s_{part}$ in the case of partial POLSAR and
  $s_{full}$ in the case of full polarimetry,
and $L$ is the number of looks.

Complex Wishart distribution statistics, however, is normally written for the scaled covariance matrix
$Z=LC_v$, whose PDF is given as:
\begin{equation}
  pdf(Z;d,\Sigma,L)=\frac{|Z|^{L-d}}{|\Sigma^L|\Gamma_d(L)}e^{-tr(\Sigma^{-1}Z)}
\end{equation}
with $\Gamma_d(L) = \pi^{d(d-1)/2} \prod^{d-1}_{i=0}\Gamma(L-i)$
and $d$ is the dimensional number of the POLSAR covariance matrix.

Our approach differs by applying the homoskedastic log transformation
  on a less-than-well-known result for the deterimant of the covarinace matrix.
  %under the assumption of circular complex Gaussian distribution
  %applicable for POLSAR data.
In \cite{Goodman_1963_AMS_178}, %Goodman
it has been proven that the ratio between the observable and the expected values of the sample covariance matrix's determinants
  behaves like a product of $d$ chi-squared random variables with different degrees of freedom 
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi^2 (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

Its log-transformed variable consequently 
  behaves like a summation of $d$ log-chi-squared random variables with the same degrees of freedom  
\begin{equation}
\Lambda^d_L = ln \left[ (2L)^d \frac{|C_v|}{|\Sigma_v|} \right] \sim \sum_{i=0}^{d-1} \Lambda^\chi (2L-2i)
\label{eqn:sum_log_chi_squared_rv}
\end{equation}
with
  $\Lambda^\chi (k) \sim \ln \left[ \chi^2 (k) \right]$

\section{Hetoroskedastic POLSAR data and the Homoskedastic Log-Transformation}
\label{sec:polsar_heterosked_model_and_log_transform}

In this section the multiplicative nature of POLSAR data is illustrated.
Log-transformation is shown converting this into a more familiar additive model.
Heteroskedasticity, which is defined as the dependence of variance upon the underlying signal,
  is proved to be the case for the original POLSAR data.
In log-transformed domain, the case for a homoskedastic model,
  where  variance is fixed and be independent of the underlying mean,
  is demonstrated.
To keep the section flowing, the mathematical derivation is only presented here in major sketches.
For more detailed derivation, Appendix \ref{chap:appendix_a} is to be reviewed.

From Eqn. \ref{eqn:prod_chi_squared_rv} and Eqn. \ref{eqn:sum_log_chi_squared_rv}
we have:
\begin{eqnarray}
  |C_v| &\sim& |\Sigma_v| \cdot \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi^2 (2L-2i) \label{eqn:determinant_distribution} \\
  \ln|C_v| &\sim& \ln|\Sigma_v| - d \cdot \ln(2L) + \sum^{d-1}_{i=0} \Lambda(2L-2i)
\label{eqn:log_determinant_distribution}  
\end{eqnarray}

In a given homogeneous POLSAR area, the parameters $\Sigma_v$, $d$ and $L$ can be considered as constant.
Thus Eqn. \ref{eqn:determinant_distribution} gives the theoretical explaination that 
  in the original POLSAR domain, a multiplicative speckle noise pattern is present.
At the same time, Eqn. \ref{eqn:log_determinant_distribution} shows that
  the logarithmic transformation converts this into a more familiar additive noise.

Since chi-squared random variables $X\ \sim\ \chi^2(k)\ $ follows the PDF:
\begin{equation}
pdf(x;k) =
  \frac{x^{(k/2)-1} e^{-x/2}}{2^{k/2} \Gamma\left(\frac{k}{2}\right)}
\label{eqn:chi_squared_dist_pdf:chap4}
\end{equation}
Applying the variable change theorem, 
  its log-transformed variable follows the PDF of:
\begin{equation}
  pdf(x;2L=k) = \frac{e^{Lx-e^x/2}}{2^{L}\Gamma(L)}
\end{equation}

Given the probability distribution function, the characteristic function of both the chi-squared and log-chi-squared random variables
  can be written as:
  \begin{eqnarray}
    CF_\chi(t) &=& (1-2it)^{−L} \\ 
    CF_\Lambda(t) &=& 2^{it} \frac{\Gamma(L+it)}{\Gamma(L)} \label{eqn:log_chi_squared_characteristic_function}
  \end{eqnarray}
Subsequently their means and variances can be computed from the given characteristic functions as:
  \begin{eqnarray}
    avg \left[ \chi(2L) \right]&=&2L \\
var \left[ \chi(2L) \right]&=&4L \\
avg \left[ \Lambda(2L) \right] &=& \psi^0(L) + \ln2 \\
var \left[ \Lambda(2L) \right] &=& \psi^1(L)
  \end{eqnarray}
  where $\psi^0()$ and $\psi^1()$ stands for digamma and trigamma function respectively.

Since the average and variance of both chi-squared distribution and log-chi-squared distribution are constant,
  the product and summation of these random variables also has fixed summary statistics.
Specifically:
\begin{eqnarray*}
  avg \left[ \prod^{d-1}_{i=0} \chi^2(2L-2i) \right] &=& 2^d \cdot \prod^{d-1}_{i=0} (L-i), \\
  var \left[ \prod^{d-1}_{i=0} \chi^2(2L-2i) \right] &=& \prod^{d-1}_{i=0} 4(L-i)(L-i+1) - \prod^{d-1}_{i=0} 4(L-i)^2, \\
  avg \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &=& d \cdot \ln{2} + \sum^{d-1}_{i=0} \psi^0(L-i), \\
  var \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &=& \sum^{d-1}_{i=0} \psi^1(L-i)
\end{eqnarray*}

Combining these results with Eqns. \ref{eqn:determinant_distribution} and \ref{eqn:log_determinant_distribution}, we have:
\begin{eqnarray}
  avg \left[ |C_v| \right]  &=& \frac{|\Sigma_v|}{L^d} \prod^{d-1}_{i=0} (L-i)\\
  var \left[ |C_v| \right]  &=&   \frac{|\Sigma_v|^2 \left[ \prod^{d-1}_{i=0} (L-i)(L-i+1) - \prod^{d-1}_{i=0} (L-i)^2 \right] }{L^{2d}} \label{eqn:var_det_is_heteroskedastic}\\
  avg \left[ \ln |C_v| \right] &=& \ln |\Sigma_v| - d \cdot \ln{L}  + \sum^{d-1}_{i=0} \psi^0(L-i) \label{eqn:avg_log_det} \\
  var \left[ \ln |C_v| \right] &=&  \sum^{d-1}_{i=0} \psi^1(L-i) \label{eqn:var_log_det_is_homoskedastic}
\end{eqnarray}

Over a real and captured image, while the paramaters $d$ and $L$ do not change for the whole image,
  the underlying $\Sigma_v$ is expected to differ from one region to the next.
Thus over an heterogeneous scene, the stochastic process for $|C_v|$ and $\ln |C_v|$ varies depending on the underlying signal $\Sigma_v$. 
  %and it is clear that 
  %$det|C_v|$ data is multiplicative in nature and its log-transformed variable $\ln (det|C_v|)$ is additive model.
In such context, Eqn. \ref{eqn:var_det_is_heteroskedastic} indicates that the variance of $|C_v|$ also differs depending on the underlying signal $\Sigma_v$, which indicates its heteroskedastic property.
At the same time, in the log-transformed domain, Eqn. \ref{eqn:var_log_det_is_homoskedastic} shows that
  the variance of $\ln |C_v|$ is invariant and independent of $\Sigma_v$ manifesting its homoskedastic nature.

\section{Consistent Measures of Distance for POLSAR data}

Similar to the way dispersion and contrast is defined in our previous work \cite{Le_2013_TGRS_SAR_MSE},
  this section introduces the consistent sense of distance in a few different perspective.
First assumming that the true value of the underlying signal $\Sigma_v$ is known \textit{a priori},
the following random variables,
  namely ratio ($\mathbb{R}$) and log-distance ($\mathbb{L}$),
  are observable according to their definitions:
%Eqns. \ref{eqn:prod_chi_squared_rv} and \ref{eqn:sum_log_chi_squared_rv} lead straight to the definition of the following random variables, which is the :

\begin{eqnarray}
  \mathbb{R} &=& \frac{|C_v|}{|\Sigma_v|} \label{eqn:determinant_ratio_observables}\\
  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}
From another perspective where the POLSAR is known coming from an homogeneous area but the true value of the underlying signal $\Sigma_v$ is \textit{unknown}, consider the dispersion ($\mathbb{D}$) and contrast ($\mathbb{C}$) random variables being defined as:
\begin{eqnarray}
  \mathbb{D} &=& \ln{|C_v|} - avg(\ln{|C_v|}) \label{eqn:dispersion_observable}\\
  \mathbb{C} &=& \ln(|C_{v1}|) - \ln(|C_{v2}|) \label{eqn:contrast_observable}
\end{eqnarray}

Taking the results from Eqns. \ref{eqn:determinant_distribution}, \ref{eqn:log_determinant_distribution} and \ref{eqn:avg_log_det} we have
\begin{eqnarray}
\mathbb{R} &\sim& \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi^2 (2L-2i) \label{eqn:determinant_ratio_distribution} \\
\mathbb{L} &\sim&  \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln(2L)
\label{eqn:log_determinant_distance_distribution} \\ 
 \mathbb{D} &\sim& \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln{2} + k
\label{eqn:dispersion_distribution} \\ 
 \mathbb{C} &\sim& \sum^{d-1}_{i=0} \Delta(2L-2i)
\label{eqn:contrast_distribution}  
\end{eqnarray}
with $\Delta(2L) \sim \Lambda(2L) - \Lambda(2L)$
and $k=\sum^{d-1}_{i=0} \psi^0(L-i)$

Also given the characteristic functions (CF) for the elementary components $\Lambda(2L)$ written in Eqn. \ref{eqn:log_chi_squared_characteristic_function}, 
  Appendix \ref{sec:appendix_b} derives the characteristic functions for the summative random variables as:
\begin{align}
  CF_{\Lambda^d_L}(t) &= \frac{2^{idt}}{\Gamma(L)^d} \prod^{d-1}_{j=0} \Gamma(L-j+it) \\
  CF_{\mathbb{L}}(t) &= \frac{1}{L^{idt} \Gamma(L)^d} \prod^{d-1}_{j=0} \Gamma(L-j+it) \\
  CF_{\mathbb{D}}(t) &= \frac{e^{ikt}}{\Gamma(L)^d} \prod^{d-1}_{j=0} \Gamma(L-j+it) \\
  CF_{\Delta(2L)} &= \frac{\Gamma(2L) B(L-it,L+it)}{\Gamma(L)^2} \\
  CF_{\mathbb{C}}(t) &=  \prod^{d-1}_{j=0} \frac{\Gamma(2L-2j) B(L-j-it,L-j+it)}{\Gamma(L-j)^2}
\end{align}

Since each elementary simulation component follows fixed distributions (i.e. $\chi^2(2L), \Lambda(2L), ... $),
  it is natural that these variables also follow fixed distributions.
Moreover, they are independent to $\Sigma_v$.
This serves as conclusive evidence that
  these random variables follows consistent and fixed distributions,
  regardless of the underlying signal $\Sigma_v$.

\section{SAR as the Special Case of Polarimetric SAR}
\label{sec:sar_special_case_of_polsar}

The prevous section has validated the use of our models for 3-dimensional ($d=3$) full polarimetric and 2-dimensional ($d=2$) partial polarimetric SAR cases.
In this section, the focus is on the limit case where the dimensional number is reduced to 1, i.e. $d=1$.
Physically this means the multi-dimensional POLSAR dataset is collapsed into the one-dimensional and classical SAR data.
Mathmatically, the sample covariance matrix is reduced to the sample variance
  and the determinant equates the scalar value.
On another hand, it is well known that for SAR data, variance equals intensity.
Thus the special case of our result is investigated carefully and is shown to be consistent with previous results for SAR intensity data.
This can be thought of 
  either as a cross-validation evidence for the proposed POLSAR models
  or alternatively as having SAR as the special case of POLSAR. 
%  with well known results for SAR intensity is presented.

The results so far for our models can be summarized as:
\begin{align}
  \mathbb{R} &= \frac{|C_v|}{|\Sigma_v|} \sim \frac{1}{(2L)^d} \prod^{d-1}_{i=0} \chi^2(2L-2i) \\% \label{eqn:polsar_ratio_det_cov_dist} \\
  \mathbb{L} &= \ln{|C_v|} - \ln{|\Sigma_v|} \sim \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln{2L} \\ %\label{eqn:polsar_dispersion_log_det_cov_dist} \\
  \mathbb{D} &= \ln{|C_v|} - avg(\ln{|C_v|}) \sim \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \ln{2} + k\\
  \mathbb{C} &= \ln{|C_{1v}|} - \ln{|C_{2v}|} \sim \sum^{d-1}_{i=0} \Delta(2L-2i) \\
  \mathbb{A} &= avg(\mathbb{L}) = \sum^{d-1}_{i=0} \psi^0(L-i) - d \cdot \ln{L} \\ %\label{eqn:polsar_dispersion_averages} \\
  \mathbb{V} &= var(\mathbb{L}) = \sum^{d-1}_{i=0} \psi^1(L-i) \\ %\label{eqn:polsar_dispersion_variance} \\
  \mathbb{E} &= mse(\mathbb{L}) =\left[ \sum^{d-1}_{i=0} \psi^0(L-i) - d \cdot \ln{L} \right]^2 +  \sum^{d-1}_{i=0} \psi^1(L-i) \label{eqn:polsar_dispersion_mse} 
\end{align}

Upon setting $d=1$ into the above models,
  Section \ref{sec:appendix_sar_special_case_of_polsar} shows that the reduced results are consistent with
not only the following results from the our previous works on single-look SAR \cite{Le_2013_TGRS_SAR_MSE}, i.e. $d=L=1$,
\begin{align*}
  I &\sim \bar{I} \cdot pdf \left[ e^{-R} \right] \\
  \log_2{I} &\sim \log_2{\bar{I}} + pdf \left[ 2^xe^{-2^x}\ln2 \right] \\
  \mathbb{R} &= \frac{I}{\bar{I}} \sim pdf \left[ e^{-x} \right]  \\
  \mathbb{L} &= \log_2{I} - \log_2{\bar{I}} \sim pdf \left[ 2^xe^{-2^x}\ln2 \right]\\
  \mathbb{D} &= \log_2{I} - avg(\log_2{I}) \sim pdf \left[ e^{-(2^xe^{-\gamma})} 2^xe^{-\gamma} \ln2 \right] \\
  \mathbb{C} &= \log_2{I_1} - \log_2{I_2} \sim pdf \left[ \frac{2^x}{(1+2^x)^2} \ln2 \right] \\
  \mathbb{A} &= avg(\mathbb{L}) = -\gamma / \ln{2} \\
  \mathbb{V} &= var(\mathbb{L}) = \frac{\pi^2}{6} \frac{1}{ \ln^2{2}} \\
  \mathbb{E} &= mse(\mathbb{L}) = \frac{1}{\ln^2{2}}( \gamma^2 + \pi^2/6 ) = 4.1161 
\end{align*}
but also the following well-known results for multi-look SAR, i.e. $d=1,L>1$:
  \begin{eqnarray}
I &\sim& pdf \left[ \frac{L^L x^{L-1} e^{-Lx/\bar{I}}}{\Gamma(L) \bar{I}^L} \right] \\
N = \ln{I} &\sim& pdf \left[ \frac{L^L}{\Gamma(L)} e^{L(x-\bar{N})-Le^{x-\bar{N}}} \right]
  \end{eqnarray}
Furthermore, the following results for multi-look SAR data,
  which can be thought of 
    either as extensions of the corresponding single-look SAR results
    or as simple cases of the POLSAR results
  are also derived as:
  \begin{align*}
    \mathbb{R} &= \frac{I}{\bar{I}} \sim pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \label{eqn:multi_look_SAR_ratio_dist} \right]\\
    \mathbb{L} &= \ln{I} - \ln{\bar{I}} \sim pdf \left[ \frac{L^Le^{Lt-Le^t}}{ \Gamma(L)}  \right] \\
    \mathbb{D} &= \ln{I} - avg(\ln{I}) \sim pdf \left[ \frac{e^{L[x-\psi^0(L)]-e^{[x-\psi^0(L)]}}}{\Gamma(L)} \right] \\
    \mathbb{C} &= \ln{I_1} - \ln{I_2} \sim pdf \left[ \frac{e^{x}}{(1+e^x)^{2}} \right] \\
    \mathbb{A} &= avg(\mathbb{L}) = \psi^0(L) - \ln{L} \\
    \mathbb{V} &= var(\mathbb{L}) = \psi^1(L) \\
    \mathbb{E} &= mse(\mathbb{L}) = \left[ \psi^0(L) - \ln{L} \right]^2 + \psi^1(L)
  \end{align*}

This newly derived models for multi-look SAR data can also be validated against real-life data.
Fig. \ref{fig:verify_multi_look_SAR_dispersion_contrast_models} presents the the results of an experiment carried out for the stated purpose.
In the experiment the intensity of a single-channel SAR data (HH) for a homogeneous area in the AIRSAR Flevoland dataset is extracted.
Then the histograms for the log-distance and and contrast is plotted against the theoretical PDF given above.
The ENL is set to the nominal number of 4.
And good visual match is apparent in the final results.

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[verify multi-look SAR log-distance]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_multi_look_sar_dispersion_pdf.eps} 	
		 \label{multi_look_dispersion}
	} 
	\hfill	
	\subfloat[verify multi-look SAR contrast]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_multi_look_sar_contrast_pdf.eps} 	
		 \label{multi_look_contrast}
	}
\end{tabular}
\caption{Multi-Look SAR dispersion and contrast: modelled response matches very well with real-life captured data.}
\label{fig:verify_multi_look_SAR_dispersion_contrast_models}
\end{figure}

\section{Validating the models against real-life data}
\label{sec:polsar_models_validation}

\subsection{Explaining practical data with the given Number-of-Looks}

This section describes an experiment to validate the models above against real-life captured data.
The validation procedure is quite straightforward.
Given that the stochastic models have been derived in the previous sections, 
  they can be graphically visualized by the histogram plots of the simulated data.
At the same time, the form of the real-life practical data is also observable via histogram plots of the data samples extracted from an homogeneous area.
Therefore, the theoretical models can be validated 
  if for the same parameters 
  the two plots match each other reasonably well.

As a demonstration, a homogeneous area was chosen from the AIRSAR Flevoland POLSAR data as experimental data samples.
Then theoretical models are employed in an attempt to explain the data.
The validating models include:
  the determinant and its log-transformed models, together with the dissimilarity measures namely: the determinant ratio, the log-distance, the dispersion and the contrast measures of distance.

The models are closely related.
Given the same parameter set, the determinant and determinant ratio are just scaled version of each other.
Meanwhile, the log-determinant, the log-distance and the dispersion are also just shifted version of each other.
Thus ideally speaking
  if one model is validated all the other models will also be,
  assuming that all the parameters of the image are known exactly.
Still in this section, all the models will be made subject to investigation.

Among all these models, the least-assumed stochastic process for dispersion and contrast measures of distance are validated first.
For each pixel in the region, the determinant of the covariance matrix is computed and the log-transformation is applied.
Then the average of POLSAR covariance matrix’s determinant in the log-transformed domain, i.e. $avg(ln|C_v|)$, is measured for dispersion.
Subsequently the observable samples of dispersion and contrast are computed according to Eqns. \ref{eqn:dispersion_observable} and \ref{eqn:contrast_observable}.
Then their histograms are traced out.

At the same time, theoretical simulations according to Eqns. \ref{eqn:dispersion_distribution} and \ref{eqn:contrast_distribution} are carried out.
The nominal value of 4 was taken as the dataset's number of look,
 while the dimensional number is set to 3 or 2 respectively depending on whether full or partial polarimetric SAR dataset is being investigated.
All the histogram plots are presented in Fig. \ref {fig:verify_polsar_2x2_simulation_dispersion_contrast}.
Apparently, a visual match is observable which verify the applicability of the theoretical models for the dispersion and contrast measures of distance.

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[part-pol (2x2) dispersion]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_2x2_dispersion_distribution.eps} 
		 \label{dispersion_2x2}
	} 
	\hfill	
	\subfloat[part-pol (2x2) contrast]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_2x2_contrast_distribution.eps} 	
		 \label{contrast_2x2}
	} \\
	\subfloat[full-pol (3x3) dispersion]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_3x3_dispersion_distribution.eps} 
		 \label{dispersion_3x3}
	} 
	\hfill	
	\subfloat[full-pol (3x3) contrast]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_3x3_contrast_distribution.eps} 	
		 \label{contrast_3x3}
	}
\end{tabular}
\caption{Validating the dispersion and contrast models against both partial and full polarimetric AIRSAR Flevoland data.}
\label{fig:verify_polsar_2x2_simulation_dispersion_contrast}
\end{figure}

Apart from dispersion and contrast,
the other four models to be investigated require an estimation of the ``true'' underlying signal $|\Sigma_v|$. 
There are two ways to estimate this quantity over an homogeneous area.
The traditional way is to simply set the true signal equal to the average of the polsar covariance matrix in its original domain, i.e. $\Sigma_v = avg(C_v$).
Another approach is to estimate the true signal from the average of the log-determinant of the polsar covariance matrix (i.e. $avg[ln|C_v|]$) using Eqn. \ref{eqn:avg_log_det}.
Both approaches will be explored in this section.
However, given that the log-determinant average has already been computed earlier, 
  the second approach is used first for the validation of determinant-ratio and log-distance.

Fig. \ref{fig:verify_polsar_2x2_simulation_det_ratio_log_distance} validate the models of determinant-ratio and log-distance against real-life data.
In this experiment, the theoretical models is simulated from Eqns \ref{eqn:determinant_ratio_distribution} and \ref{eqn:log_determinant_distance_distribution},
  while the observable samples are computed using Eqns \ref{eqn:determinant_ratio_observables} and \ref{eqn:log_distance_observables}
  with the true signal estimated from the log-determinant average, i.e. $avg(\ln|C_v|)$.
Again a reasonable match is observed which validates the models for log-distance and determinant ratio.

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[part-pol (2x2) determinant ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_2x2_determinant_ratio_distribution.eps} 
		 \label{determinant_ratio_2x2}
	} 
	\hfill	
	\subfloat[part-pol (2x2) log distance]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_2x2_log_distance_distribution.eps} 	
		 \label{log_distance_2x2}
	} \\
	\subfloat[full-pol 3x3 determinant ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_3x3_determinant_ratio_distribution.eps} 
		 \label{determinant_ratio_3x3}
	} 
	\hfill	
	\subfloat[full-pol 3x3 log distance]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_3x3_log_distance_distribution.eps} 	
		 \label{log_distance_3x3}
	}
\end{tabular}
\caption{Validating determinant-ratio and log-distance models with $|\Sigma_v|$ is computed using $avg(\ln|C_v|)$}
\label{fig:verify_polsar_2x2_simulation_det_ratio_log_distance}
\end{figure}

Since the models for the determinant and log-determinant are just scaled or shifted version of the models for determinant-ratio and log-distance, similar validation results are to be expected. 
And if the  true signal are computed in the same manner as described before then in fact similar match can be easily observed.

However, a more interesting  phenomena is to be described.
It happens in the validation process for determinant and its log-transformed model,
  where the theoretical response is taken from the simulated stochastic process described by Eqns. \ref{eqn:determinant_distribution} and \ref{eqn:log_determinant_distribution}.
%For such simulation, an estimation of  the true signal is needed.
The interesting phenomena happens when the true signal is estimated by the first approach i.e. equal to the average of the sample covariance matrix in its original domain.
Interestingly, this approach results in a different estimation for the true signal with reference to the method described earlier.

Subsequently the validation plots, which are presented in Fig \ref{fig:verify_polsar_2x2_simulation_det}, exhibit some small discrepancies.
The differences are easier to observe in the log-determinant plots, 
  where except for some small shifted effects, the shapes of the models and the practical data exhibits certain resemblance.

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[verify polsar 2x2: determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_2x2_determinant_distribution.eps} 
		 \label{determinant_2x2}
	} 
	\hfill	
	\subfloat[verify polsar 2x2: log-determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_2x2_log_det_distribution.eps} 	
		 \label{log_det_2x2}
	} \\ 
	\subfloat[verify polsar 3x3: determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_3x3_determinant_distribution.eps} 
		 \label{determinant_3x3}
	} 
	\hfill	
	\subfloat[verify polsar 3x3: log-determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_3x3_log_det_distribution.eps} 	
		 \label{log_det_3x3}
	} 
\end{tabular}
\caption{Validating determinant and log-determinant models with $\Sigma_v = avg(C_v)$}
\label{fig:verify_polsar_2x2_simulation_det}
\end{figure}

In short the least-assumed dispersion and contrast measures of distance are shown to match reasonably well with the practical data.
The same can be stated for the other four models, namely: determinant, log-determinant, determinant ratio and log-distance,
  if the underlying parameters can be estimated reasonably well for the given image.   
However as described above a single ``true signal'' $|\Sigma_v|$ can have two different estimated values,
  depending on which estimation method was being used.
The discrepancy suggests that at least one parameter for the models was wrongly employed.

But what model parameter were used wrongly, and even if that can be corrected, would a better match become observable?
The question is answered in Section \ref{sec:improve_the_match_bw_theory_practice}, 
  where not only the look number is shown to be misused
  but also the match of between the theoretical model and the practical data is shown to improve as well once a better look number (ENL) is estimated.
For now, let us contend that:
  using appropriate estimation of the parameters, 
  the proposed models match reasonably well with the practical data.

\subsection{Comparing Theoretical Assumptions and Practical Implementations}
\label{sec:improve_the_match_bw_theory_practice}

Even though the initial assumptions for the proposed theory is intentionally kept to minimal, 
  like all other similar models, the proposed model in this paper is built upon certain presumptions.
Practical conditions however may not always satisfy these prerequisites.
In this section, certain gaps between the conditions found in practical real-life data and the theoretical assumptions are discussed.
And it is shown that: the theoretical model proposed can handle the practical data, even when these ``imperfections''  are taken into account.

There are two main ``imperfections'' that are usually found in practical POLSAR data with reference to the theoretical model.
The first is the mutually independent assumption for each component in the POLSAR target vector $s$.
Practically however high correlation is routinely observable between the POLSAR data components,
  specifically between $S_{hh}$ and $S_{vv}$.
This phenomena also presents in our AIRSAR dataset, where $\Sigma_v = avg(C_v) = \begin{vmatrix} 0.0084 & 1 \cdot 10^{-6} + 4 \cdot 10^{-4} i & 0.0071 - 0.0017 i \\ 1 \cdot 10^{-6} - 4 \cdot 10^{-4} i & 0.0017 & -3 \cdot 10^{-4} - 2 \cdot 10^{-4} i \\ 0.0071 + 0.0017 i & -3 \cdot 10^{-4} + 2 \cdot 10^{-4} i & 0.0122 \end{vmatrix}$.
Despite the mismatch,
  astute reader would probably notice that the proposed model apparently still valid under such condition as evidenced by the part-pol (HH-VV) and full-pol plots in Figs. \ref{fig:verify_polsar_2x2_simulation_dispersion_contrast} to \ref{fig:verify_polsar_2x2_simulation_det}.
This suggests that the proposed model is also applicable on correlated POLSAR components,
  eventhough a full explaination for this, however, is outside the scope of this paper.

Another assumption of the model (for both SAR and POLSAR) is that the samples are statistically independent to each other.
This is a reasonable assumption given that 
  the transmission and receipt of analog signals are done independently for each radar pulse, i.e. for each resolution cell.
Thus theoretically speaking, adjacent pixels in an image can be assumed to be statistically independent.

However, the actual imaging mechanism of a real-life (POL)SAR processor is that of digital nature,
where the analog signal is to be converted into a digital data-set. 
Specifically, the analog signal in SAR 
  which is characterized by the pulse bandwidth measurement,
  is fed into an analog-to-digital sampling and conversion process 
  which is characterized by its sampling rate.
Theoretically it is possible to define a sampling rate to ensure that each digital pixel correspond exactly to an analog physical cell resolution.
Practically however, to ensure ``perfect reconstruction'' of the analog signal, the sampling rate is normally set at a slightly higher value than the theoretical mark.
This results in a higher number of samples / pixels than the number of physical cells available in the scene.

Stated differently, in practical data each physical radar cell is spread to more than one pixel
  and each pixel now contains less than one physical analog cell.
This high sampling rate also results in 
  a significantly higher correlation between pairs of pixels that maybe related within a single physical cell resolution 
  than the correlation found between pairs of pixels that are further away and hence having less physical relation to each other.
It also results in reduced effective number of look, 
  in which say a window of 3x3 pixel actually contains less than 9 physical analog cells. 
The former phenomena is partially explained in \cite{Raney_1988_TGRS_666} for SAR,
  while the later is experimental observed for POLSAR data in \cite{Lee_1994_TGRS_1017} and \cite{Anfinsen_2009_TGRS_3795}.
The oversampling practice is also documented by the producers of SAR processors.
For AIRSAR, the sampling rate and pulse bandwidth combinations are either 90/40MHz or 45/20MHz \cite{JPL_2013_Web_AIRSAR_Impl}.
While for RadarSat2, the pixel resolution and range - azimuth resolutions for SLC fine-quad mode is advertised as $(4.7 \cdot 5.1)m^2/(5.2 \cdot 7.7)m^2$ \cite{MDA_2013_Web_RadatSat2_Description}

The proposed model can handle this imperfection that is available in pratical data.
The trick is that instead of using the nominal Number of Look given by the SAR processor,
  an ENL estimation procedure is employed. %the model validation can make use of the Effective Number of Looks that is manually estimated from a given practical dataset.
The first part of this section details a simple ENL estimation technique for POLSAR data.
While the second part demonstrates how the practical imperfection manifest itself and how it can be handled in a RADARSAT2 dataset.
It also illustrates how the match shown in Section \ref{sec:polsar_models_validation} for the AIRSAR Flevoland dataset can be improved,
  by using ENL estimation.

\subsection{Estimating the Effect Number-of-Looks (ENL)}

>>>>

\appendix

\chapter{Log-Chi-Square Distribution}
\label{chap:appendix_a}
